---
title: "Model Interpretability"
author: "Jeremiah W"
date: "December 2020"
output: html_document
---

<style type="text/css">
  body{
  font-size: 18pt;
}
</style>

# Resources
[Vid](https://youtu.be/eNvKnhMJd2o)
[DataDic](https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data)
[IML vignette](https://cran.r-project.org/web/packages/iml/vignettes/intro.html)
[PFI](https://christophm.github.io/interpretable-ml-book/feature-importance.html#advantages-9)
[ALE](https://christophm.github.io/interpretable-ml-book/ale.html)
[LIME](https://christophm.github.io/interpretable-ml-book/lime.html)
[R txbk](https://bradleyboehmke.github.io/HOML/iml.html)

```{r setup, include=FALSE}
#options(digits = 2, scipen = 2)
library(tidyverse)
library(tidymodels)
library(AmesHousing)
library(broom)
library(iml);library(pdp)
a = make_ames()
```

```{r}
a = a %>%
  select_if(is.numeric) %>% 
  relocate(Sale_Price)
```

```{r warning=FALSE, message=FALSE}
lm = lm(Sale_Price ~ . , data = a)
lm %>% tidy
```

#Regression w/effect plots

### problem: 'coef/estimate plots'
```{r warning=FALSE, message=FALSE, fig.width=12}
#coef plot w/confi intervals
lm %>% tidy(conf.int = TRUE) %>%  #AWESOME, can get conf intervals w/this arg
  .[-1,] %>% 
  ggplot(aes(x = term, y = estimate)) +
  geom_point() +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  coord_flip() +
  scale_y_log10()
```

<PROBLEM> the problem with this plot is that the features are not in the same units of measure/scales 

e.g.
LotFrontage: Linear FEET of street connected to property
<VS>
FullBath: COUNT Full bathrooms above grade

### solution: 'effects plot'
```{r fig.width=12}
#multiple coefs by actual values to see *contribution*
a %>% mutate(pk = row_number()) %>%
  pivot_longer(cols = c(everything(), -Sale_Price, -pk)) %>% left_join(
    lm %>% tidy %>% 
    slice(-1)  %>%  #don't need intercept since it's a constant 
    select(term, estimate),
    by = c('name' = 'term')
  ) %>% mutate(
    effect = value * estimate,
    pos_neg = if_else(effect < 0, 'Negative', 'Positive'),
    effect = abs(effect)
  ) %>% ggplot(aes(name, effect, color = pos_neg)) +
  geom_boxplot() + coord_flip() +
  scale_y_log10()

```
# Model Agnostic Methods

### 1 Create and Fit Models
```{r}
### tidymodels:spec models

rf = rand_forest() %>% 
  set_mode('regression') %>% 
  set_engine('ranger')

xgb = boost_tree() %>% 
  set_mode('regression') %>% 
  set_engine('xgboost')

### tidymodels:fit models

set.seed(345)
rf = rf %>% fit(Sale_Price ~ ., data = a)
xgb = xgb %>% fit(Sale_Price ~ ., data = a)
```

### 2 Create Predictor Obj Components
[Ref](http://uc-r.github.io/iml-pkg#procedures)
```{r}
# 1. create a data frame with just the vars/features/x
x.data = a %>% select(-Sale_Price) %>% as.data.frame()

# 2. create vector of the outcome/y
y.data = a$Sale_Price
```

### 3 Create Predictor Obj
```{r}
# Similar to DALEX and lime, the *Predictor* object holds the (1) model, (2) the features data, and the (3) outcome data

rf.predictor = Predictor$new(rf, data = x.data, y = y.data)

xgb.predictor = Predictor$new(xgb, data = x.data, y = y.data)
```

## PFI: *FeatureImp*
Permutation Feature Importance Plots

The feature importance measure works by calculating the 'loss' in performance measure after permuting the feature. 

### Create & Viz FeatureImp Obj
[Ref](http://uc-r.github.io/iml-pkg#vip)
```{r message=FALSE}
(rf.fimp = FeatureImp$new(rf.predictor, loss = 'mae'))
(xgb.fimp = FeatureImp$new(xgb.predictor, loss = 'mae'))

rf.fimp %>% plot
xgb.fimp %>% plot
```
Interpretation: If Gr_Liv_Area is removed, model loses of its predictive power

## ALE: *FeatureEffects*
Accumulated Local Effects

Shows distribution of average effects from each feature

ALE plots are generally 'better' than pdp plots since they account for feature correlation.  PDP plots assume features are uncorrelated, which is rarely the case IRL (see proof features ARE correleated below)

Predictor > FeatureEffects > plot
Predictor > FeatureEffect > plot #NOT plural

### Check: Confirm Feature Correlation 
```{r fig.width=12}
a %>% DataExplorer::plot_correlation()
a %>% dlookr::correlate() %>% filter(coef_corr > 0.5 | coef_corr < -0.5) %>% arrange(-coef_corr)
```

### Create & Plot FeatureEffects object
```{r fig.width=15, fig.height=15, message=FALSE, warning=FALSE}
#marginal effect of each feature
rf.ale = FeatureEffects$new(rf.predictor)
xgb.ale = FeatureEffects$new(xgb.predictor)

rf.ale %>% plot
xgb.ale %>% plot
```

[PDP vs ALE](https://christophm.github.io/interpretable-ml-book/ale.html#motivation-and-intuition)

To summarize how each type of plot (PDP, ALE) calculates the effect of a feature at a certain grid value v:
<br><br>
Partial Dependence Plots: "Let me show you what the model predicts on average when each data instance has the value v for that feature. I *ignore whether the value v makes sense for all data instances*."
<br><br>
ALE plots: "Let me show you how the model predictions change in a small "window" of the feature around v *for data instances* in that window."

## *TreeSurrogate* (Global Surrogate Model)
```{r}
kidai = 木代ｗ

rf.kidai = TreeSurrogate$new(rf.predictor, maxdepth = 3)
xgb.kidai = TreeSurrogate$new(xgb.predictor, maxdepth = 3)

rf.kidai %>% plot
xgb.kidai %>% plot

#making predictions using the surrogate model
cbind(
rf.kidai$predict(a) %>% head,
xgb.kidai$predict(a) %>% head
) %>% rename(
  rf = '.y.hat'
) %>% rename(
  xgb = '.y.hat'
)
```


## LIME *LocalModel* (Local Surrogate Model)

### find the observation with the greatest Gr_Liv_Area
```{r}
#GrLivArea: Above grade (ground) living area square feet
x.data %>% pull(Gr_Liv_Area) %>% sort(decreasing = TRUE) %>% .[1]
x.data %>% mutate(rn = row_number()) %>% filter(Gr_Liv_Area == 5642) %>% relocate(rn, Gr_Liv_Area)
```

### explain the prediction for that observation
```{r}
(rf.lime = LocalModel$new(rf.predictor, x.interest = x.data[1499,]))
(xgb.lime = LocalModel$new(xgb.predictor, x.interest = x.data[1499,]))

rf.lime %>% plot
xgb.lime %>% plot
```




## *Interaction*
<h style="color: red; font-size:18px;">Warning: Takes ~10m to run</h>
```{r eval=FALSE}
rf.interact = Interaction$new(rf.predictor)
rf.interact %>% plot
```
