---
title: "Predicting Hotel Type and Booking Demand"
author: 'Jeremiah Wang'
date: "November 2020"
output:
  html_document:
    theme: simplex
    highlight: pygments
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
    df_print: paged
---

```{r}
#https://datatables.net/reference/option/
options(DT.options = list(scrollX = TRUE, pagin=TRUE, fixedHeader = TRUE, searchHighlight = TRUE))
```

```{r include=FALSE}
library(DataExplorer);library(data.table);library(dlookr);
library(extrafont);library(formattable);library(GGally);library(here);
library(janitor);library(lubridate);library(naniar);
library(PerformanceAnalytics);
library(plotly);library(RColorBrewer);library(readxl);
library(skimr);library(tidyverse);library(scales)

library(tidymodels);library(h2o)
```

# Introduction

[Check out this Kaggle](https://www.kaggle.com/jessemostipak/hotel-booking-demand)


# Get Data

```{r message=FALSE}
a = read_csv('hotel_bookings.csv') %>%
  clean_names() %>% 
  mutate(across(where(is.character), factor)) %>% 
  select(sort(tidyselect::peek_vars())) %>% 
  select(
    where(is.Date),
    where(is.factor),
    where(is.numeric)
  )
```

# Resources

1. adr
    + https://www.xotels.com/en/glossary/adr-average-daily-rate/
2. transient
    + https://www.xotels.com/en/glossary/transient/
3. group rate
    + https://www.xotels.com/en/glossary/group-rate/
4. Guide to Hotel Management
    + https://www.4hoteliers.com/features/article/9886
5. distribution channels
    + https://www.xotels.com/en/glossary/channels/

## 5 min EDA

```{r}
a %>% head
```

```{r}
skimr::skim(a)
```

### clean data
```{r}
# these numeric vars s/b factor vars
a = a %>% mutate_at(vars(arrival_date_day_of_month, arrival_date_week_number, arrival_date_year, is_canceled, is_repeated_guest), factor)

# reordering df
a = a %>% select(sort(tidyselect::peek_vars())) %>% 
  select(
    where(is.Date),
    where(is.factor),
    where(is.numeric)
  )
```


# EDA: datatime vars

### range
```{r}
a %>% select(where(is.Date)) %>% colnames()
paste(
  'The date range of this dataset is from',
  a %>% pull(reservation_status_date) %>% range %>% .[1],
  'to',
  a %>% pull(reservation_status_date) %>% range %>% .[2]
)
```




### time series count graph -- ungrouped
```{r message=FALSE}
a %>% count(reservation_status_date, name = 'count') %>%
  arrange(reservation_status_date) %>%
  plot_ly(
    x = ~reservation_status_date,
    y = ~count
  ) %>% layout(
    title = 'booking counts by date',
    xaxis = list(title = ''),
    yaxis = list(title = '')
    )
```

### time series count graph -- grouped (hotel)
```{r message=FALSE}
a %>% group_by(reservation_status_date, hotel) %>%
  summarise(count = n()) %>% 
  arrange(reservation_status_date) %>%
  plot_ly(
    x = ~reservation_status_date,
    y = ~count,
    color = ~hotel,
    alpha = 0.7
  ) %>% layout(
    title = 'booking counts by date/hotel',
    xaxis = list(title = ''),
    yaxis = list(title = '')
    )
```

### time series count graph -- grouped (customer_type)
```{r message=FALSE}
a %>% group_by(reservation_status_date, customer_type) %>%
  summarise(count = n()) %>% 
  arrange(reservation_status_date) %>%
  plot_ly(
    x = ~reservation_status_date,
    y = ~count,
    color = ~customer_type,
    alpha = 0.7
  ) %>% layout(
    title = 'booking counts by date/customer_type',
    xaxis = list(title = ''),
    yaxis = list(title = '')
    )
```

### time series count graph -- grouped (deposit_type)
```{r message=FALSE}
a %>% group_by(reservation_status_date, deposit_type) %>%
  summarise(count = n()) %>% 
  arrange(reservation_status_date) %>%
  plot_ly(
    x = ~reservation_status_date,
    y = ~count,
    color = ~deposit_type,
    alpha = 0.7
  ) %>% layout(
    title = 'booking counts by date/deposit_type',
    xaxis = list(title = ''),
    yaxis = list(title = '')
    )
```

### time series count graph -- grouped (distribution_channel)
```{r message=FALSE}
a %>% group_by(reservation_status_date, distribution_channel) %>%
  summarise(count = n()) %>% 
  arrange(reservation_status_date) %>%
  plot_ly(
    x = ~reservation_status_date,
    y = ~count,
    color = ~distribution_channel,
    alpha = 0.7
  ) %>% layout(
    title = 'booking counts by date/distribution_channel',
    xaxis = list(title = ''),
    yaxis = list(title = '')
    )
```

# EDA: nom vars

### sample data
```{r}
a %>% select(where(is.factor)) %>% slice_sample(n = 10)
```

### glimpse structure
```{r}
a %>% select(where(is.factor)) %>% glimpse
```

### check missing values
```{r}
a %>% select(where(is.factor)) %>% miss_var_summary
```

### distribution of level counts per factor
```{r}
jpal = colorRampPalette(brewer.pal(8,'Dark2'))(15)

a %>% select(where(is.factor)) %>%
  map(n_unique) %>%
  as.tibble() %>%
  pivot_longer(everything()) %>%
  plot_ly(y = ~name, x = ~value, color = ~name, colors = jpal) %>%
  add_bars() %>%
  hide_legend() %>% 
  layout(
    title = 'distribution of level counts per factor',
    xaxis = list(title = ''),
    yaxis = list(title = '')
    )
```

### reference: names of unique levels
```{r}
a %>% select(where(is.factor)) %>%
  map(unique)

a = a %>% mutate(arrival_date_month = factor(arrival_date_month, levels = c('January','February','March','April','May','June','July','August','September','October','November','December')))
```

# EDA: num vars

### check missing values
```{r}
a %>% select(where(is.numeric)) %>% miss_var_summary
```

### sample data
```{r}
a %>% select(where(is.numeric)) %>% slice_sample(n = 10)
```

### glimpse structure
```{r}
a %>% select(where(is.numeric)) %>% glimpse
```

### viz: distribution histogram
```{r}
a %>% select(where(is.numeric)) %>% DataExplorer::plot_histogram(nrow = 2, ncol = 1)
```

### viz: distribution density
```{r}
a %>% select(where(is.numeric)) %>% DataExplorer::plot_density(nrow = 2, ncol = 1)
```

### viz: distribution bivariate
```{r}
a %>% select(hotel, where(is.numeric)) %>% DataExplorer::plot_boxplot(by = 'hotel', nrow = 3, ncol = 1)
```

### ADR further investigation
```{r}
#looks like one outlier in adr is changing the scale, making it hard to see the true distribution --remove outlier
a$adr %>% range
a %>% filter(adr != 5400) %>%
  select(hotel, adr) %>%
  plot_ly(y = ~hotel, x = ~adr, color = ~hotel) %>% 
  add_boxplot()

#https://stackoverflow.com/questions/57300053/split-a-plotly-boxplot-x-axis-by-group
a %>% filter(adr != 5400) %>%
  select(hotel, adr, customer_type) %>%
  plot_ly(y = ~hotel, x = ~adr, color = ~customer_type, colors = jpal, group = ~customer_type) %>% 
  add_boxplot() %>% 
  layout(
    boxmode = 'group', #SUPER IMPORTANT
    title = 'ADR by Hotel/customer_type'
    ) 

#https://stackoverflow.com/questions/57300053/split-a-plotly-boxplot-x-axis-by-group
a %>% filter(adr != 5400) %>%
  select(hotel, adr, market_segment) %>%
  plot_ly(y = ~hotel, x = ~adr, color = ~market_segment, colors = jpal, group = ~market_segment) %>% 
  add_boxplot() %>% 
  layout(
    boxmode = 'group', #SUPER IMPORTANT
    title = 'ADR by Hotel/market_segment'
    ) 

#https://stackoverflow.com/questions/57300053/split-a-plotly-boxplot-x-axis-by-group
a %>% filter(adr != 5400) %>%
  select(hotel, adr, arrival_date_month) %>%
  plot_ly(y = ~hotel, x = ~adr, color = ~arrival_date_month, colors = jpal, group = ~arrival_date_month) %>% 
  add_boxplot() %>% 
  layout(
    boxmode = 'group', #SUPER IMPORTANT
    title = 'ADR by Hotel/arrival_date_month'
    ) 

```

```{r}


  

  
a %>% filter(adr != 5400) %>%
  group_by(hotel, adr, customer_type) %>%
  do(
    list.of.plots = plot_ly(. , x = ~adr, y = ~hotel) %>% 
      add_boxplot(
        color = ~customer_type,
        )
  ) %>% subplot(nrows = 2, shareY = TRUE, shareX = TRUE)   
  

```


```{r}
library(DescTools)

ggplotly(
a %>% filter(adr != 5400) %>% 
  mutate(arrival_date_month = factor(arrival_date_month, labels = DescTools::StrLeft(levels(a$arrival_date_month), 3))) %>% 
  group_by(arrival_date_month, hotel) %>% 
  summarise(med.adr = median(adr, na.rm = TRUE)) %>% 
  ggplot(aes(arrival_date_month, med.adr, fill = hotel)) +
  geom_col(position = 'dodge') +
  scale_fill_manual(values = c('blue4','darkorange'))
) %>% layout(
  title = 'Median ADR by Hotel/Month'
)
  
```

### correlations: viz
```{r}
a %>% select(where(is.numeric)) %>% dlookr::plot_correlate()
a %>% select(where(is.numeric)) %>% GGally::ggcorr(palette = "RdBu", label = TRUE)
```

