---
title: "Housing Prices"
subtitle: "Predicting Housing Prices using RF, ENET, XGBST"
author: 'Jeremiah Wang'
date: "December 2020"
output:
  html_document:
    theme: simplex
    highlight: pygments
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
    df_print: paged
---

```{r}
#https://datatables.net/reference/option/
options(DT.options = list(scrollX = TRUE, pagin=TRUE, fixedHeader = TRUE, searchHighlight = TRUE))
```

```{r include=FALSE}
library(DataExplorer);library(data.table);library(dlookr);
library(extrafont);library(formattable);library(funModeling);library(GGally);
library(here);library(janitor);library(lubridate);library(naniar);
library(PerformanceAnalytics);library(plotly);library(RColorBrewer);
library(readxl);library(skimr);library(tidyverse);library(scales);
library(visdat)
```

# Intro

[Check this Kaggle](https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data)

# Data Dic
```{r}
SalePrice - the property's sale price in dollars. This is the target variable that you're trying to predict.
1stFlrSF: First Floor square feet
2ndFlrSF: Second floor square feet
3SsnPorch: Three season porch area in square feet
Alley: Type of alley access
Bedroom: Number of bedrooms above basement level
BldgType: Type of dwelling
BsmtCond: General condition of the basement
BsmtExposure: Walkout or garden level basement walls
BsmtFinSF1: Type 1 finished square feet
BsmtFinSF2: Type 2 finished square feet
BsmtFinType1: Quality of basement finished area
BsmtFinType2: Quality of second finished area (if present)
BsmtFullBath: Basement full bathrooms
BsmtHalfBath: Basement half bathrooms
BsmtQual: Height of the basement
BsmtUnfSF: Unfinished square feet of basement area
CentralAir: Central air conditioning
Condition1: Proximity to main road or railroad
Condition2: Proximity to main road or railroad (if a second is present)
Electrical: Electrical system
EnclosedPorch: Enclosed porch area in square feet
ExterCond: Present condition of the material on the exterior
Exterior1st: Exterior covering on house
Exterior2nd: Exterior covering on house (if more than one material)
ExterQual: Exterior material quality
Fence: Fence quality
FireplaceQu: Fireplace quality
Fireplaces: Number of fireplaces
Foundation: Type of foundation
FullBath: Full bathrooms above grade
Functional: Home functionality rating
GarageArea: Size of garage in square feet
GarageCars: Size of garage in car capacity
GarageCond: Garage condition
GarageFinish: Interior finish of the garage
GarageQual: Garage quality
GarageType: Garage location
GarageYrBlt: Year garage was built
GrLivArea: Above grade (ground) living area square feet
HalfBath: Half baths above grade
Heating: Type of heating
HeatingQC: Heating quality and condition
HouseStyle: Style of dwelling
Kitchen: Number of kitchens
KitchenQual: Kitchen quality
LandContour: Flatness of the property
LandSlope: Slope of property
LotArea: Lot size in square feet
LotConfig: Lot configuration
LotFrontage: Linear feet of street connected to property
LotShape: General shape of property
LowQualFinSF: Low quality finished square feet (all floors)
MasVnrArea: Masonry veneer area in square feet
MasVnrType: Masonry veneer type
MiscFeature: Miscellaneous feature not covered in other categories
MiscVal: $Value of miscellaneous feature
MoSold: Month Sold
MSSubClass: The building class
MSZoning: The general zoning classification
Neighborhood: Physical locations within Ames city limits
OpenPorchSF: Open porch area in square feet
OverallCond: Overall condition rating
OverallQual: Overall material and finish quality
PavedDrive: Paved driveway
PoolArea: Pool area in square feet
PoolQC: Pool quality
RoofMatl: Roof material
RoofStyle: Type of roof
SaleCondition: Condition of sale
SaleType: Type of sale
ScreenPorch: Screen porch area in square feet
Street: Type of road access
TotalBsmtSF: Total square feet of basement area
TotRmsAbvGrd: Total rooms above grade (does not include bathrooms)
Utilities: Type of utilities available
WoodDeckSF: Wood deck area in square feet
YearBuilt: Original construction date
YearRemodAdd: Remodel date
YrSold: Year Sold
```


# Get & Split Data
```{r}
a = read_csv('train.csv') %>% 
  clean_names() %>% 
  mutate(across(where(is.character), factor)) %>% 
  select(sort(tidyselect::peek_vars())) %>% 
  select(
    where(is.Date),
    where(is.factor),
    where(is_character),
    where(is_numeric)
  )

a %>% sample_n(10)
```

```{r}
mo_sold, year_built, year_remod_add, yr_sold
```


# 5 min EDA

### viz missing
```{r fig.width=12, fig.height=9}
total.cols = a %>% colnames() %>% length() 

a %>% visdat::vis_dat()
```

### check missing, na, inf
```{r}
a %>% select(where(is.factor)) %>% funModeling::status() %>% arrange(-q_na)
a %>% select(where(is.numeric)) %>% funModeling::status() %>% arrange(-q_na)
```

Observations

1. 4 factor vars have 80%+ missing values
    + since this is so many, consider removing entirely rather than imputing
2. For other factor vars with minimal missing nas, consider either:
    + replacing missing values with dedicated 'unknown' level OR
    + replacing missing values by imputing
3. Consider imputing num vars with nas
    + lot_frontage: rf or kmeans imputation with all other vars
    + garage_yr_blt: rf imputation with other date vars (e.g. year_XXX, yr_sold)
    + mas_vnr_area: rf or kmeans imputation with all other vars


# Wrangling / Cleaning
```{r eval=FALSE}
a$pool_qc = NULL
a$misc_feature = NULL
a$alley = NULL
a$fence = NULL
```

# EDA: nom vars

### names
```{r}
(nom.vars = a %>% select(where(is.factor)) %>% colnames %>% as.character)
```

### check missing
```{r}
a %>% select(nom.vars) %>% miss_var_summary
```

### sample
```{r}
a %>% select(nom.vars) %>% sample_n(5)
```

### skim
```{r}
a %>% select(nom.vars) %>% skim_without_charts()
```

### viz: level counts distribution
```{r}
a %>% select(nom.vars) %>% map(n_unique) %>% as.data.frame.list %>% pivot_longer(everything()) %>% mutate(name = fct_reorder(name, value)) %>% plot_ly(x = ~value, y = ~name, color = ~name) %>% hide_legend() %>% layout(title = 'factor vars level counts', xaxis = list(title = 'count'), yaxis = list(title = ''))

a %>% select(nom.vars) %>% funModeling::freq()
```

### viz: mosaic plots: ggpairs
```{r fig.width=12, fig.height=9}
#documentation: https://mran.microsoft.com/snapshot/2016-01-12/web/packages/GGally/vignettes/ggpairs.html

a %>% select(nom.vars) %>% 
  GGally::ggpairs(
    columns = c(nom.vars[13], nom.vars[32]),
    mapping = aes(color = eval(as.name(nom.vars[13])))
)

```

Description of Masonry veneer type

+ BrkCmn	Brick Common
+ BrkFace	Brick Face
+ CBlock	Cinder Block
+ None	None
+ Stone	Stone

Description of Exterior Quality Levels

+ Ex	Excellent
+ Gd	Good
+ TA	Average/Typical
+ Fa	Fair
+ Po	Poor

<h style="color: blue; font-size:16px;">Not too surprisingly, Many of the Brick type Masonry veneer type had Exterior Qualities of Good and Excellent</h>

# EDA: num vars

### names
```{r}
(num.vars = a %>% select(where(is.numeric)) %>% colnames %>% as.character)
```

### check missing
```{r}
a %>% select(num.vars) %>% miss_var_summary
```

### sample
```{r}
a %>% select(num.vars) %>% sample_n(5)
```

### skim
```{r}
a %>% select(num.vars) %>% skim_without_charts()
```

### viz: distribution - hist
```{r}
a %>% select(num.vars) %>% DataExplorer::plot_histogram(
  scale_x = 'log10',
  #geom_histogram_args = list(bins = 50L),
  ncol = 2, nrow = 2
)
```

### viz: distribution - density
```{r}
a %>% select(num.vars) %>% DataExplorer::plot_density(
  scale_x = 'log10',
  #geom_histogram_args = list(bins = 50L),
  ncol = 2, nrow = 2
)
```

### viz: outliers
```{r}
a %>% select(num.vars) %>% dlookr::plot_outlier()
```

### viz: normality
```{r}
a %>% select(num.vars) %>% dlookr::plot_normality()
```
### viz: normality: outcome var only
```{r}
a %>% select(sale_price) %>% dlookr::plot_normality()
```
<h style="color: blue; font-size:16px;">consider log transforming</h>

### viz: correlations

```{r fig.width= 12, fig.height=9}
a %>% select(num.vars) %>% visdat::vis_cor()

a %>% select(num.vars) %>% dlookr::plot_correlate()

a %>% select(num.vars) %>% GGally::ggcorr(low = 'red', high = 'darkgreen', label = TRUE)
```

Observations:

1. Not surprisingly there are positive cor among certain vars (list below not comprehensive):
    + gr_liv_area & tot_rms_abv_grd
    + gr_liv_area & sale_price
    + bedroom_abv_gr & tot_rms_abv_grd
    + yr_blt & garage_yr_blt
2. Consider doing PCA on correlated vars

# EDA: nom/num vars

```{r}
target.var = 'sale_price'
```
### viz: cor w/target - pearson
```{r}
a %>% select(num.vars) %>% funModeling::correlation_table(target = target.var)
```

<h style="color: blue; font-size:16px;">quick and dirty pseudo feature importance</h>


### feature engineering a binary var for viz below
```{r}
a %>% count(overall_qual) %>% arrange(-overall_qual)

a = a %>% mutate(
  overall_qual.gte.8.flg = if_else(
    overall_qual >= 8, 'yes', 'no'
  )
)

```


### viz: cross plot - input/target distribution
```{r}
### useful when outcome var is a binary var
a %>% funModeling::cross_plot(
  input = nom.vars,
  target = "overall_qual.gte.8.flg"
  )
```

### quick analysis: binary target
```{r}
#This function is used to analyze data when we need to reduce variable cardinality in predictive modeling.
#works in conjunction with 'cross_plot';
a %>% funModeling::categ_analysis(
    input = nom.vars[13],
    target = 'overall_qual.gte.8.flg'
  )
```

<h style="color: blue; font-size:16px;">Of the subset of homes whose exterior quality was excellent, 88.5% had an overall quality rating of 8 or better </h>

#gomi

1. remove vars
2. impute missing
3. log transform outcome