---
title: "Power Plants - Predicting Power"
author: 'Jeremiah W'
date: "November 2020"
output:
  rmdformats::readthedown:
    code_folding: hide
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# MISC

```{r eval=FALSE}
# regarding 'readthedown' theme
# https://cran.r-project.org/web/packages/rmdformats/vignettes/introduction.html
```

<style>
    body .main-container {
        max-width: 1000px;
    }
</style>

```{r include=FALSE}
#https://datatables.net/reference/option/
options(DT.options = list(scrollX = TRUE, pagin=TRUE, fixedHeader = TRUE, searchHighlight = TRUE))

library(DataExplorer);library(data.table);library(dlookr);
library(extrafont);library(formattable);library(GGally);library(here);
library(janitor);library(lubridate);library(naniar);
library(PerformanceAnalytics);
library(plotly);library(RColorBrewer);library(readxl);
library(skimr);library(tidyverse);library(scales)

library(caret);library(tidymodels);library(prophet)

library(rmdformats);library(tufte)
```

# Introduction

[Check out this Kaggle](https://www.kaggle.com/anikannal/solar-power-generation-data)

*This data has been gathered at two solar power plants in India over a 34 day period. It has two pairs of files - each pair has one power generation dataset and one sensor readings dataset. The power generation datasets are gathered at the inverter level - each inverter has multiple lines of solar panels attached to it. The sensor data is gathered at a plant level - single array of sensors optimally placed at the plant.*

## Business Need

**Can we predict the power generation for next couple of days? This will help with better grid management**

## Statistical Method

* facebook time-series PROPHET modeling
    1. at the aggregate plant level (all inverters)
    2. at the individual inverter level, and then rollup (aggregate sum) to power plant level**
    
# Power Generation Inverter data set: Plant 1

## Get and Clean Data

```{r message=FALSE}

p1.gd = read_csv('Plant_1_Generation_Data.csv') %>%
  slice_sample(prop = 1) %>% #!!<NOTE>temp, working with a sample of datset for speed purposes
  clean_names() %>%  #lowercase
  select(sort(tidyselect::peek_vars())) %>% #sort cols alphabetically
  select(where(is.factor),where(is.character),where(is.numeric)) #sort cols by data type

#OlsonNames()
#https://stackoverflow.com/questions/41479008/what-is-the-correct-tz-database-time-zone-for-india

p1.gd = p1.gd %>% mutate(
  date_time = as.POSIXct(strptime(p1.gd$date_time, "%d-%m-%Y %H:%M"), tz = 'Asia/Kolkata'),
  source_key = factor(p1.gd$source_key)
) %>% rename(inverter = source_key) %>% 
  arrange(date_time)

p1.gd$plant_id = NULL

```

### sample rows

```{r}
p1.gd %>% slice_sample(n = 10) %>% DT::datatable()
```

### glimpse structure
```{r}
p1.gd %>% glimpse()
```

### check missing values
```{r}
p1.gd %>% miss_var_summary()
```

## EDA

### counts of each factor's unique levels

```{r}
sapply(p1.gd %>% select(where(is.factor)), n_unique) %>% as.data.frame()
```

### reference: names of unique levels
```{r}
sapply(p1.gd %>% select(where(is.factor)), unique) %>% as.data.frame() %>% arrange()
```

### time series/anomoly Plot

<h style="color: blue; font-size:12px;">**for better visibility, right click and open in new tab**</h>

```{r warning=FALSE, message=FALSE, fig.width=7, fig.height= 35, cache=TRUE, eval=TRUE}
library(anomalize)
# anomalize(data, target, method = c("iqr", "gesd"), alpha = 0.05, max_anoms = 0.2, verbose = FALSE)

# alpha: Controls the width of the "normal" range. Lower values are more conservative while higher values are less prone to incorrectly classifying "normal" observations.
# max_anoms: The maximum percent of anomalies permitted to be identified.

p1.gd.anomalize = p1.gd %>% arrange(date_time) %>% 
  mutate(inverter = fct_reorder(inverter, -daily_yield)) %>% 
  group_by(inverter) %>%
  time_decompose(daily_yield, method = 'twitter', merge = TRUE) %>%
  anomalize(remainder, alpha = 0.05, method = 'gesd') %>%
  time_recompose()

  p1.gd.anomalize %>%
    plot_anomalies(
      ncol = 1,
      alpha_dots = 0.5,
      alpha_circles = 0.5,
      size_circles = 1.5,
      time_recomposed = TRUE,
      alpha_ribbon = 0.05
      ) + scale_y_continuous(labels = comma) +
    labs(x = '', y = 'daily_yield', title = 'daily_yield')
```

<h style="color: red; font-size:18px;">**bvBOhCH3iADSZry** still has a few outliers after June 01, but its overall individual contribution shouldn't mess up aggregate predictions'</h>

### finalizing Data Set
```{r eval=FALSE}
p1.gd = p1.gd %>% filter(date_time >= '2020-06-01 00:00:00 IST')
```
<h style="color: blue; font-size:12px;">NOTE: Initially, I had filtered the data to use just clean history with no anomalies, but ultimately decided against this for two reasons.  First, I wanted to give as much history as possible since the total history is relatively short.  Second, prophet does a fabulous job of ignoring outliers'</h>

# Forecasting at the Aggregate Level

### aggregating total power at the daily level
```{r}
p1.agg = p1.gd %>% group_by(date_time) %>% summarise(daily_yield = sum(daily_yield), .groups = 'drop')
p1.agg %>% plot_ly(x = ~date_time, y = ~daily_yield) %>% add_lines() %>% layout(title = 'Plant 1 total daily_yield KW')

```

### Create Model
```{r fig.width = 12, message=FALSE}
#renaming cols to prophet's col conventions
p1.agg = p1.agg %>% select(ds = date_time, y = daily_yield)

#creating model
p1.agg.model = p1.agg %>% prophet()

#using model make future period df
p1.agg.future = p1.agg.model %>% make_future_dataframe(
  periods = 96 * 7, #there are 96 fifteen mintue periods in one day, so this represents 7 days
  freq = 15 * 60) # data is at the 15 minute frequency level (freq is in seconds)

#make forecasts df
p1.agg.forecast = p1.agg.model %>% predict(p1.agg.future)

#plot forecast
p1.agg.model %>% dyplot.prophet(p1.agg.forecast)

#plot forecast components
p1.agg.model %>% prophet_plot_components(p1.agg.forecast)

```

```{r}

```


# Power Generation Inverter data set: Plant 2

## Get and Clean Data

```{r message=FALSE}

p2.gd = read_csv('Plant_2_Generation_Data.csv') %>%
  slice_sample(prop = 1) %>% #!!<NOTE>temp, working with a sample of datset for speed purposes
  clean_names() %>%  #lowercase
  select(sort(tidyselect::peek_vars())) %>% #sort cols alphabetically
  select(where(is.POSIXct), where(is.factor),where(is.character),where(is.numeric)) #sort cols by data type

#OlsonNames()
#https://stackoverflow.com/questions/41479008/what-is-the-correct-tz-database-time-zone-for-india

p2.gd = p2.gd %>% mutate(
  #date_time = as.POSIXct(strptime(p2.gd$date_time, "%d-%m-%Y %H:%M"), tz = 'Asia/Kolkata'),
  source_key = factor(p2.gd$source_key)
) %>% rename(inverter = source_key) %>% 
  arrange(date_time)

p2.gd$plant_id = NULL

```

### sample rows

```{r}
p2.gd %>% slice_sample(n = 10) %>% DT::datatable()
```

### glimpse structure
```{r}
p2.gd %>% glimpse()
```

### check missing values
```{r}
p2.gd %>% miss_var_summary()
```

## EDA

### counts of each factor's unique levels

```{r}
sapply(p2.gd %>% select(where(is.factor)), n_unique) %>% as.data.frame()
```

### reference: names of unique levels
```{r}
sapply(p2.gd %>% select(where(is.factor)), unique) %>% as.data.frame() %>% arrange()
```

### time series/anomoly Plot

<h style="color: blue; font-size:12px;">**for better visibility, right click and open in new tab**</h>

```{r warning=FALSE, message=FALSE, fig.width=7, fig.height= 35, cache=TRUE, eval=TRUE}
library(anomalize)
# anomalize(data, target, method = c("iqr", "gesd"), alpha = 0.05, max_anoms = 0.2, verbose = FALSE)

# alpha: Controls the width of the "normal" range. Lower values are more conservative while higher values are less prone to incorrectly classifying "normal" observations.
# max_anoms: The maximum percent of anomalies permitted to be identified.

p2.gd.anomalize = p2.gd %>% arrange(date_time) %>% 
  mutate(inverter = fct_reorder(inverter, -daily_yield)) %>% 
  group_by(inverter) %>%
  time_decompose(daily_yield, method = 'twitter', merge = TRUE) %>%
  anomalize(remainder, alpha = 0.05, method = 'gesd') %>%
  time_recompose()

  p2.gd.anomalize %>%
    plot_anomalies(
      ncol = 1,
      alpha_dots = 0.5,
      alpha_circles = 0.5,
      size_circles = 1.5,
      time_recomposed = TRUE,
      alpha_ribbon = 0.05
      ) + scale_y_continuous(labels = comma) +
    labs(x = '', y = 'daily_yield', title = 'daily_yield')
```

<h style="color: red; font-size:18px;">It's a good idea to calculate the aggregate excluding the 4 inverters that were not operating from late May to end of May.  We should forecast them separately'</h>

### identify the 4 faulty inverters
```{r}
p2.gd %>% count(inverter) %>% arrange(n) %>% head(4)
p2.faulty.inverters = p2.gd %>% count(inverter) %>% arrange(n) %>% head(4) %>% select(inverter) %>% pull %>% as.character

```

1. 'good' inverters (NON-faulty inverters that didn't have complete shutdowns)
2. 'bad' inverters (FAULTY inverters that had complete shutdowns)


### finalizing Data Sets
```{r}
#filter to only good inverters
p2.gd.good = p2.gd %>% filter(!inverter %in% p2.faulty.inverters)

#filter to only bad inverters
p2.gd.bad = p2.gd %>% filter(inverter %in% p2.faulty.inverters)
```
<h style="color: blue; font-size:12px;">NOTE: Initially, I had filtered the data to use just clean history with no anomalies, but ultimately decided against this for two reasons.  First, I wanted to give as much history as possible since the total history is relatively short.  Second, prophet does a fabulous job of ignoring outliers'</h>

# Forecasting at the Aggregate Level for 'good' inverters

### aggregating total power at the daily level
```{r}
p2.agg.good = p2.gd.good %>% group_by(date_time) %>% summarise(daily_yield = sum(daily_yield), .groups = 'drop')
p2.agg.good %>% plot_ly(x = ~date_time, y = ~daily_yield, color = I('black')) %>% add_lines() %>% layout(title = 'Plant 2 good inverters total daily_yield KW')

```

### Create Model
```{r}
#renaming cols to prophet's col conventions
p2.agg.good = p2.agg.good %>% select(ds = date_time, y = daily_yield)

#creating model
p2.agg.good.model = p2.agg.good %>% prophet()

#using model make future period df
p2.agg.good.future = p2.agg.good.model %>% make_future_dataframe(
  periods = 96 * 7, #there are 96 fifteen mintue periods in one day, so this represents 7 days
  freq = 15 * 60,
  include_history = TRUE
  ) # data is at the 15 minute frequency level (freq is in seconds)

#make forecasts df
p2.agg.good.forecast = p2.agg.good.model %>% predict(p2.agg.good.future)

p2.agg.good.forecast %>% head %>% DT::datatable()

#plot forecast
p2.agg.good.model %>% dyplot.prophet(p2.agg.good.forecast)

#plot forecast components
p2.agg.good.model %>% prophet_plot_components(p2.agg.good.forecast)

```

# Forecasting at the Aggregate Level for 'bad' inverters

### aggregating total power at the daily level
```{r}
p2.agg.bad = p2.gd.bad %>% group_by(date_time) %>% summarise(daily_yield = sum(daily_yield), .groups = 'drop')
p2.agg.bad %>% plot_ly(x = ~date_time, y = ~daily_yield, color = I('darkorange')) %>% add_lines() %>% layout(title = 'Plant 2 bad inverters total daily_yield KW')
```

<h style="color: blue; font-size:12px;">For the aggregate bad inverter dataset (p2.agg.bad) ONLY, let's exclude the 'breakdown period' and provide only history from May 31st when training the model</h>

### Create Model
```{r message=FALSE}
#renaming cols to prophet's col conventions
p2.agg.bad = p2.agg.bad %>% filter(date_time >= '2020-05-31 UTC') %>% select(ds = date_time, y = daily_yield) 

#creating model
p2.agg.bad.model = p2.agg.bad %>% prophet()

#using model make future period df
p2.agg.bad.future = p2.agg.bad.model %>% make_future_dataframe(
  periods = 96 * 7, #there are 96 fifteen mintue periods in one day, so this represents 7 days
  freq = 15 * 60,
  include_history = TRUE
  ) # data is at the 15 minute frequency level (freq is in seconds)

#make forecasts df
p2.agg.bad.forecast = p2.agg.bad.model %>% predict(p2.agg.bad.future)

p2.agg.bad.forecast %>% head %>% DT::datatable()

#plot forecast
p2.agg.bad.model %>% dyplot.prophet(p2.agg.bad.forecast) %>% dygraphs::dyOptions(colors = c('darkorange','blue'))

#plot forecast components
p2.agg.bad.model %>% prophet_plot_components(p2.agg.bad.forecast)

```

```{r}

```

