---
title: "Credit Card Customer Clustering"
author: 'Jeremiah W'
date: "October 29, 2020"
output:
  html_document:
    theme: simplex
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
    df_print: paged
---

```{r}
#https://datatables.net/reference/option/
options(DT.options = list(scrollX = TRUE, pagin=TRUE, fixedHeader = TRUE, searchHighlight = TRUE))
```

```{r include=FALSE}
library(DataExplorer);library(data.table);library(dlookr)
library(extrafont);library(formattable);library(GGally);library(here);
library(janitor);library(lubridate);library(naniar);
library(patchwork);library(PerformanceAnalytics);
library(plotly);library(RColorBrewer);library(readxl);
library(skimr);library(tidyverse);library(scales)

library(NbClust);library(factoextra);library(ape);library(dbscan)
```

# Introduction
[Please read this Kaggle](https://www.kaggle.com/arjunbhasin2013/ccdata)

# Data Dictionary
```{r eval=FALSE}
# CUSTID : Identification of Credit Card holder (Categorical)
# 
# BALANCE : Balance amount left in their account to make purchases (
# 
# BALANCEFREQUENCY : How frequently the Balance is updated, score between 0 and 1 (1 = frequently updated, 0 = not frequently updated)
# 
# PURCHASES : Amount of purchases made from account
# 
# ONEOFFPURCHASES : Maximum purchase amount done in one-go
# 
# INSTALLMENTSPURCHASES : Amount of purchase done in installment
# 
# CASHADVANCE : Cash in advance given by the user
# 
# PURCHASESFREQUENCY : How frequently the Purchases are being made, score between 0 and 1 (1 = frequently purchased, 0 = not frequently purchased)
# 
# ONEOFFPURCHASESFREQUENCY : How frequently Purchases are happening in one-go (1 = frequently purchased, 0 = not frequently purchased)
# 
# PURCHASESINSTALLMENTSFREQUENCY : How frequently purchases in installments are being done (1 = frequently done, 0 = not frequently done)
# 
# CASHADVANCEFREQUENCY : How frequently the cash in advance being paid
# 
# CASHADVANCETRX : Number of Transactions made with "Cash in Advanced"
# 
# PURCHASESTRX : Numbe of purchase transactions made
# 
# CREDITLIMIT : Limit of Credit Card for user
# 
# PAYMENTS : Amount of Payment done by user
# 
# MINIMUM_PAYMENTS : Minimum amount of payments made by user
# 
# PRCFULLPAYMENT : Percent of full payment paid by user
# 
# TENURE : Tenure of credit card service for user
```


In one piped statement:
  
1. read in data
2. rename all colnames lowercase
3. convert char to factor vars
4. order cols by name: alphabetically
5. order cols by datatype: nominal, then numeric

# Get Data
```{r message=FALSE}
a = read_csv('CreditCard_Customers.csv') %>% 
  clean_names() %>% 
  mutate(across(where(is.character),factor)) %>% 
  select(sort(tidyselect::peek_vars())) %>% 
  select(where(is.factor),where(is.numeric))
```
### quick skim
```{r}
a %>% head %>% DT::datatable()
a %>% glimpse
```
<h4 style="color: blue; font-size:12px;">There appears to be outliers for male annual income</h4>
<h4 style="color: blue; font-size:12px;">XXX</h4>

### remove unecessary col
```{r}
a = a %>% select(-cust_id)
```

# EDA: num vars

### check missing values
```{r}
miss_var_summary(a)
```


### distribution: viz
```{r fig.width=12, fig.height=9}
DataExplorer::plot_histogram(a, nrow = 3, ncol = 3)
```

[dlookr vignette](https://github.com/choonghyunryu/dlookr)

### outliers: viz
```{r}
#find names of cols with outliers_ratio > 10%
(a %>% diagnose_outlier() %>% filter(outliers_ratio > 5))
outlier.cols = a %>% diagnose_outlier() %>% filter(outliers_ratio > 5) %>% select(variables) %>% pull()

a %>% plot_outlier(all_of(outlier.cols))
```

### normality: viz
```{r}
a %>% plot_normality(
  colnames(a)[1],
  colnames(a)[2],
  colnames(a)[3],
  colnames(a)[4],
  colnames(a)[5],
  colnames(a)[6],
  colnames(a)[7],
  colnames(a)[8],
  colnames(a)[9],
  colnames(a)[10],
  colnames(a)[11],
  colnames(a)[12],
  colnames(a)[13],
  colnames(a)[14],
  colnames(a)[15],
  colnames(a)[16],
  colnames(a)[17]
  )


```


### correlations: viz
```{r fig.width=12, fig.height=9}
dlookr::plot_correlate(a)
```


### pairwise correlations: viz
```{r fig.width=12, fig.height=9}
a %>% GGally::ggcorr(low = '#990000', mid = '#E0E0E0', high = '#009900', label = TRUE)
```

# Preprocessing
### remove missing rows
```{r}
a = a %>% na.omit()
```

### transformation: normalize data
```{r}
a.scale = a %>% scale
```

### create matrix
```{r}
a.matrix = a.scale %>% as.matrix
```

# Determine Optimal Number of Clusters

[Reference](https://rstudio-pubs-static.s3.amazonaws.com/455393_f20bacf1329a49dab40eb393308b33eb.html#choosing-optimal-number-of-clusters)

### silhouette analysis with kmeans and euclidean distancing
```{r cache=TRUE}
factoextra::fviz_nbclust(
  a.matrix,
  diss = dist(a.matrix, method = "euclidean"),
  FUNcluster=kmeans,
  method="silhouette"
  ) +
  theme_classic()
```

<h4 style="color: blue; font-size:12px;">While 2 is what is suggested, it likely makes little business sense to only have 2 clusters; will go with 3, 4, or 5</h4>

# Clustering via Hierarchial Methods
[Reference 1](http://www.sthda.com/english/wiki/beautiful-dendrogram-visualizations-in-r-5-must-known-methods-unsupervised-machine-learning)
[Reference 2](https://sites.ualberta.ca/~lkgray/uploads/7/3/6/2/7362679/slides-clusteranalysis.pdf)

```{r fig.width= 12, fig.height = 9, cache=TRUE}
dist.a = dist(a, method = 'euclidean') #Distance Matrix Computation

#ward.D2, creates groups such that variance is minimized within clusters, dissimilarities are squared before clustering
hclust.a = hclust(dist.a, method = 'ward.D2') #agglomeration method, how to group points
#hclust.a = hclust(del.distanced, method = 'complete') 

plot(hclust.a, hang = -1, cex = 0.8, main = NULL, xlab = NULL)

plot(hclust.a, hang = -1, cex = 0.8, main = NULL, xlab = NULL);abline(h = 220000, col = 'blue', lty = 2) #4 clusters at this height

#At a given height cutoff of x
#Clusters below height x  have a [dist.method] [hclust.method] distance <= height cutoff

hclust.colors = RColorBrewer::brewer.pal(4, 'Paired') #choose # of cols corresponding an optimal number of clusters

hclust.clusters = cutree(hclust.a, 4)

plot(as.phylo(hclust.a), type = 'fan', tip.color = hclust.colors[hclust.clusters],
     label.offset = 2, cex = 1.0)

plot(ape::as.phylo(hclust.a), type = 'unrooted', tip.color = hclust.colors[hclust.clusters],
     label.offset = 2, cex = 1.0, no.margin = TRUE)

```

<h4 style="color: blue; font-size:12px;">Clusters below a height cutoff of 210000  have a kmeans euclidean distance <= height cutoff</h4>
<h4 style="color: blue; font-size:12px;">Looking at the dendogram, 9 clusters might be 'over-clustering'; will stick with 5 clusters</h4>

# Clustering via Centroid Based Methods: Kmeans

## Kmeans
```{r fig.width=12, fig.height=9, cache=TRUE}
km = eclust(
  a.matrix,
  FUNcluster="kmeans",
  k=4,
  hc_metric = "euclidean",
  hc_method = "ward.D2",
  )
```

# Clustering via Centroid Based Methods: PAM

### Create Distance Matrix
```{r}
#gower distances allow you to work with both nominal and numeric data
gower_dist <- cluster::daisy(a.scale, metric = "gower")
gower_mat <- as.matrix(gower_dist)
```


### Implementing K Medoids Clustering
```{r cache=TRUE}
#Using PAM (partitioning around medoids) clustering, calculate silhouette width for k-medoids 
max.clusters = 10

sil_width <- c(NA)
  
  for(i in 2:max.clusters){
    
    pam_fit <- cluster::pam(gower_dist,
                            diss = TRUE,
                            k = i)
    
    sil_width[i] <- pam_fit$silinfo$avg.width
    
  }

gc() #clear memory
```

### Plot results
```{r fig.width=12, fig.height=9}

plot(1:10, sil_width,
       xlab = "Number of medoids/clusters",
       ylab = "Silhouette Width")
lines(1:10, sil_width)
```


### Finalize PAM model with optimal number of medoids
```{r finalfit}
pam_fit <- cluster::pam(gower_dist,
                          diss = TRUE,
                          k = 4)
```


### Visualize the Clusters
```{r fig.width=16, fig.height=12}
tsne_obj <- Rtsne::Rtsne(gower_dist, is_distance = TRUE)

tsne_data <- tsne_obj$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering))

(graph.clusters = ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster)))

```

# Summarize Clustering Results

### summary: data
```{r message=FALSE}
a = a %>% mutate(cluster = factor(km$cluster, levels = 1:4))

a %>%
  group_by(cluster) %>%
  summarise(
      mean.balance = mean(balance, na.rm = TRUE),
      mean.balance_frequency = mean(balance_frequency, na.rm = TRUE),
      mean.cash_advance = mean(cash_advance, na.rm = TRUE),
      mean.cash_advance_frequency = mean(cash_advance_frequency, na.rm = TRUE),
      mean.cash_advance_trx = mean(cash_advance_trx, na.rm = TRUE),
      mean.credit_limit = mean(credit_limit, na.rm = TRUE),
      mean.installments_purchases = mean(installments_purchases, na.rm = TRUE),
      mean.minimum_payments = mean(minimum_payments, na.rm = TRUE),
      mean.oneoff_purchases = mean(oneoff_purchases, na.rm = TRUE),
      mean.oneoff_purchases_frequency = mean(oneoff_purchases_frequency, na.rm = TRUE),
      mean.payments = mean(payments, na.rm = TRUE),
      mean.prc_full_payment = mean(prc_full_payment, na.rm = TRUE),
      mean.purchases = mean(purchases, na.rm = TRUE),
      mean.purchases_frequency = mean(purchases_frequency, na.rm = TRUE),
      mean.purchases_installments_frequency = mean(purchases_installments_frequency, na.rm = TRUE),
      mean.purchases_trx = mean(purchases_trx, na.rm = TRUE),
      mean.tenure = mean(tenure, na.rm = TRUE),
      count = n()
  )
```

### summary: viz
```{r fig.width=12, fig.height=9}
library(ggthemes)
DataExplorer::plot_boxplot(a, by = 'cluster', nrow = 3, ncol = 1)
```

### summary: viz
```{r warning=FALSE, fig.width=12, fig.height=9, cache=TRUE}

xvars = a %>% select(-cluster) %>% names %>% as.character()

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[2])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[2],' by cluster'), xaxis = list(title = xvars[2]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[3])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[3],' by cluster'), xaxis = list(title = xvars[3]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[4])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[4],' by cluster'), xaxis = list(title = xvars[4]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[5])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[5],' by cluster'), xaxis = list(title = xvars[5]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[6])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[6],' by cluster'), xaxis = list(title = xvars[6]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[7])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[7],' by cluster'), xaxis = list(title = xvars[7]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[8])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[8],' by cluster'), xaxis = list(title = xvars[8]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[9])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[9],' by cluster'), xaxis = list(title = xvars[9]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[10])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[10],' by cluster'), xaxis = list(title = xvars[10]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[11])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[11],' by cluster'), xaxis = list(title = xvars[11]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[12])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[12],' by cluster'), xaxis = list(title = xvars[12]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[13])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[13],' by cluster'), xaxis = list(title = xvars[13]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[14])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[14],' by cluster'), xaxis = list(title = xvars[14]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[15])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[15],' by cluster'), xaxis = list(title = xvars[15]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[16])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[16],' by cluster'), xaxis = list(title = xvars[16]))

a %>% plot_ly(y = ~cluster, x = ~eval(as.name(xvars[17])), color = ~cluster, colors = hclust.colors) %>% add_boxplot() %>% hide_legend() %>% layout(
  title = paste0(xvars[17],' by cluster'), xaxis = list(title = xvars[17]))

```

```{r warning=FALSE}

```


# Notes

1. Better outlier handling
2. Add elbow method for determining optimal number of clusters
3. feature scaling -- perhaps log scale would be a good idea for some of the features that are grossly NOT normally distributed
