---
title: "Exploring and Predicting Crime in Washington DC"
subtitle: "Data History from Jan 1, 2012 to November 2, 2017"
author: 'Jeremiah W, November 2020'
output:
  html_document:
    theme: cerulean
    highlight: pygments
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
    df_print: paged
---

```{r include=FALSE}
#https://datatables.net/reference/option/
options(DT.options = list(scrollX = TRUE, pagin=TRUE, fixedHeader = TRUE, searchHighlight = TRUE))

library(DataExplorer);library(data.table);library(dlookr);
library(extrafont);library(formattable);library(GGally);library(here);
library(janitor);library(lubridate);library(naniar);
library(PerformanceAnalytics);
library(plotly);library(RColorBrewer);library(readxl);
library(skimr);library(tidyverse);library(scales)
```

# Intro
[Check out this Kaggle](https://www.kaggle.com/vinchinzu/dc-metro-crime-data)

# Objective

* Understand crime in WDC by exploring data at both macro and micro levels, make recommendations for 2018, and then make future predictions.

# Goals
1. Find total offenses by each factor group
    + Historically
    + Latest Year 2017
2. Find total offenses by each offense type/ward group
    + Historically
    + Latest Year 2017
3. Trend over time (last 3 years)
    + total offenses
    + total offenses by type
4. Predict total offenses for the next 3 months

# Get Data
```{r warning=FALSE, message=FALSE}
a = read_csv('raw.csv') %>% 
  clean_names() %>% 
  select(sort(tidyselect::peek_vars())) %>% 
  select(
    report_dat,
    month, day, year, hour, minute, second,
    where(is.Date),
    where(is.character),
    where(is.factor),
    where(is.numeric)
  )
```

# Resources

1. WDC Wards
    + https://dc.curbed.com/maps/dc-council-eight-wards-map-elections-neighborhoods
2. psa = police service area
    + https://mpdc.dc.gov/page/police-districts-and-police-service-areas

### sample data
```{r}
a %>% sample_n(10)
```
### glimpse structure
```{r}
a %>% glimpse
```

### counts of unique levels
```{r}
a %>% map_df(n_unique)
```

### clean data
```{r eval=FALSE}
#remove unused cols
a = a %>% select(!c(minute, second, anc, block, block_group, end_date, ew, neighborhood_cluster, ns, start_date, voting_precinct, ccn, district, x, x1)) %>%
  mutate(
    report_dat = anytime::anydate(report_dat),
    #start_date = anytime::anydate(start_date),
    #end_date = anytime::anydate(end_date),
    across(where(is.character), factor),
    census_tract = factor(census_tract, levels = a$census_tract %>% unique %>% sort),
    ward = factor(ward, levels = a$ward %>% unique %>% sort),
    psa = factor(psa, levels = a$psa %>% unique %>% sort),
    year = factor(year, levels = a$year %>% unique %>% sort),
    month = factor(month, levels = a$month %>% unique %>% sort),
    day = factor(day, levels = a$day %>% unique %>% sort),
    hour = factor(hour, levels = a$hour %>% unique %>% sort)
    ) %>%
  select(sort(tidyselect::peek_vars())) %>% 
  select(
    where(is.Date),
    month, day, year, hour,
    where(is.character),
    where(is.factor),
    where(is.numeric) 
    ) %>% arrange(report_dat, crimetype, offense)
    
#abak = a
#saveRDS(abak, 'cleaned_data.RDS')
#a = abak
```

<h style="color: blue; font-size:12px;">If we wanted to do more detailed geographic analysis, we might want to include block and block group</h>

```{r include=FALSE}
a = readRDS('cleaned_data.RDS')
#filter to ~5 years of data
a = a %>% filter(year %in% c(2012,2013,2014,2015,2016,2017))
```

# EDA: factor vars

### sample data
```{r}
a %>% select(where(is.factor)) %>% sample_n(10)
```

### glimpse structure
```{r}
a %>% select(where(is.factor)) %>% glimpse
```

### check missing values
```{r}
a %>% select(where(is.factor)) %>% miss_var_summary
```

<h style="color: blue; font-size:12px;">With so little data missing, I feel comfortable omitting rows with NAs</h>
```{r}
a = a %>% na.omit()
```


### distribution of level counts per factor
```{r out.width = 1000, warning=FALSE}
#defining custom color palette
jpal = grDevices::colorRampPalette(brewer.pal(8,'Dark2'))(25)

a %>% select(where(is.factor)) %>% map_df(n_unique)

a %>% select(where(is.factor)) %>%
  map_df(n_unique) %>%
  pivot_longer(. , everything(), 'features') %>%
  plot_ly(x = ~features, y = ~value, color = ~features, colors = jpal) %>% 
  add_bars() %>%
  layout(title = 'Unique Level Counts per Factor')
  
```

### reference: names of unique levels
```{r}
a %>% select(where(is.factor)) %>% map(unique)
```

## Goal 1. Find total offenses by each factor group X

### Historically (2012 - 2016)
```{r fig.width= 10}
a %>% filter(year != 2017) %>% select(where(is.factor)) %>% DataExplorer::plot_bar(ncol = 1, nrow = 1, title = 'Total Offenses by Category - Historic')
```
```{r}
a %>% filter(year != 2017) %>% count(psa, sort = TRUE, name = 'count') %>%
  head(10) %>%
  mutate(psa = factor(psa)) %>%
  mutate(psa = fct_reorder(psa, -count)) %>% 
  plot_ly(x = ~psa, y = ~count, color = ~psa, colors = jpal) %>% 
  add_bars() %>% 
  layout(
    title = 'Police Service Areas with the Most Crime'
  )
```


Observations:

1. Most offenses occurs late summer to early fall (top 3 months: Oct, July, Aug)
2. More offenses occurs towards the latter half the month (>= Day 16)
3. Non-violent offenses occur ~5x as much as violent offenses
4. Theft is, by far, the most common offense
    + auto-theft is so predominant, it has its own category
5. The Northeast quad has significantly more crime
    + can partly be attributed to its slightly larger size as seen below in the quick map
6. Wards 2 and 6 are the most dangerous wards by a good margin
7. PSA service areas 207, 302, and 208 have the most crime by a good margin


### Latest Year 2017
```{r fig.width= 10}
a %>% filter(year == 2017) %>% select(where(is.factor)) %>% DataExplorer::plot_bar(ncol = 1, nrow = 1, title = 'Total Offenses by Category - 2017')
```
```{r}
a %>% filter(year == 2017) %>% count(psa, sort = TRUE, name = 'count') %>%
  head(10) %>%
  mutate(psa = factor(psa)) %>%
  mutate(psa = fct_reorder(psa, -count)) %>% 
  plot_ly(x = ~psa, y = ~count, color = ~psa, colors = jpal) %>% 
  add_bars() %>% 
  layout(
    title = 'Police Service Areas with the Most Crime'
  )
```

### quick map
```{r out.width = 700, out.height = 500}
ipal = grDevices::colorRampPalette(brewer.pal(12,'Paired'))(56)

a %>% filter(year == 2017) %>% plot_ly(x = ~xblock, y = ~yblock, color = ~quad, colors = ipal) %>% add_markers() %>% layout(title = 'WDC Quadrants')

a %>% filter(year == 2017) %>% plot_ly(x = ~xblock, y = ~yblock, color = ~ward, colors = ipal) %>% add_markers() %>% layout(title = 'WDC Wards')

a %>% filter(year == 2017) %>% plot_ly(x = ~xblock, y = ~yblock, color = ~psa, colors = ipal) %>% add_markers() %>% layout(title = 'WDC PSAs') %>% hide_legend()
```

Observations relative to Historic

1. **January and May 2017** were atypically more dangerous 
2. Non-violent offenses made up a greater percent of total offenses then they did historically (5x), occurring about ~6.5 x as much as violent offenses
3. Wards 2 and 6 continue to be the top 2 dangerous wards
4. PSA service areas 207, 302, and 208 continue to have the most crime, albeit by a smaller margin

## Recommendations Batch 1

Recommendations for the year 2018 based on 2017 findings

1. Staff More Policemen and women from 1 to 3 pm..  
2. Staff More Policemen and women in wards 2 and 6.
3. Staff More Policemen and women in PSA 208.

## Goal 2. Find total offenses by each offense type/ward group

### Historically (2012 - 2016)
```{r out.width = 1000, warning=FALSE}
ggplotly(
  a %>% filter(year != 2017) %>%
    count(ward, offense) %>% ggplot(aes(x = offense, y = n, fill = offense)) +
    geom_col() +
    coord_flip() +
    labs(x = '', y = 'count', title = 'Total Offenses by Type/Ward - Historic') +
    facet_wrap(~ward) +
    theme(legend.position = 'none')
)
```

Observations:

1. Ward 2 has the greatest number of cases as observed above, but this is disproportionately **mostly composed of theft/other offenses**
    + the offense count of its second largest offense category, theft/auto, is even slightly less than that of Ward 1's'
2. Ward 7 has a relatively high proportion of motor vehicle theft offenses    
3. Ward 8 has a relatively high proportion of burglary offenses

### Latest Year 2017
```{r out.width = 1000}
ggplotly(
  a %>% filter(year == 2017) %>% 
    count(ward, offense) %>% ggplot(aes(x = offense, y = n, fill = offense)) +
    geom_col() +
    coord_flip() +
    labs(x = '', y = 'count', title = 'Total Offenses by Type/Ward - 2017') +
    facet_wrap(~ward) +
    theme(legend.position = 'none')
)
```

Observations relative to Historic

1. Ward 2 continues to have the greatest number of cases as observed above, but this is disproportionately **mostly composed of theft/other offenses**
2. Ward 6's share of theft/other is higher
3. *Relative to other wards*, in 2017, Wards 5 and 6 are more dangerous than they were historically
    + on an *absolute basis*, all wards are less dangerous
        + see time series graphs below

## Recommendations Batch 2

Recommendations for the year 2018 based on 2017 findings

1. Work on educating citizens on proper theft deterrent methods, especially in ward 2; perhaps install more cameras there.
2. Looking at wards 7 and 8, it seems like motor vehicle theft may be correlated by assault with a dangerous weapon.
		+ Check out if both offenses occurred simultaneously.

## Goal 3. Trend over time

```{r}
# create col for start of month (a 'month' col) used for grouping and graphing
a = a %>% mutate(
  monthkey = lubridate::make_datetime(
    as.numeric(as.character(year)),
    as.numeric(as.character(month)),
    1)
  ) %>% relocate(report_dat, monthkey, everything())

```


#Total Offenses over time

### Total Offenses by Month
```{r out.width = 1000, message=FALSE}
a %>% group_by(monthkey) %>%
  summarise(total.offenses = n()) %>%
  ungroup() %>% 
  plot_ly(x = ~monthkey, y =~total.offenses) %>%
  add_lines(size = I(3)) %>% layout(
  xaxis = list(title = ''),
  yaxis = list(title = ''),
  title = 'Total Offenses by Month'
  )
```

Observations

1. Crime follows clear seasonality
    + peaks occurring late summer/early fall
2. Peaked in mid 2014
3. Troughed in early 2015

### Total Offenses by Month/Type
```{r out.width = 1000, message=FALSE}
ggplotly(
  a %>% group_by(monthkey, offense)%>%
    summarise(total.offenses = n()) %>%
    ungroup() %>%
    mutate(offense = fct_reorder(offense, total.offenses, .fun = mean)) %>% 
    ggplot(aes(monthkey, total.offenses, fill = offense)) +
    geom_area() +
    labs(title = 'Total Offenses Percentage (#) by Month/Offense Type', x = '', y = '')
)

ggplotly(
  a %>% group_by(monthkey, offense)%>%
    summarise(total.offenses = n()) %>%
    mutate(total.offenses.pct = total.offenses/sum(total.offenses)) %>% 
    ungroup() %>% 
    mutate(offense = fct_reorder(offense, total.offenses, .fun = mean)) %>% 
    ggplot(aes(monthkey, total.offenses.pct, fill = offense)) +
    geom_area() +
    scale_y_continuous(labels = scales::percent) +
    labs(title = 'Total Offenses Percentage (%) by Month/Offense Type', x = '', y = '')
)
```

### Total Offenses by Month/Ward
```{r out.width = 1000, message=FALSE}
ggplotly(
  a %>% group_by(monthkey, ward)%>%
    summarise(total.offenses = n()) %>%
    ungroup() %>%
    mutate(ward = fct_reorder(ward, total.offenses, .fun = mean)) %>% 
    ggplot(aes(monthkey, total.offenses, fill = ward)) +
    geom_area() +
    labs(title = 'Total Offenses Percentage (#) by Month/Ward', x = '', y = '')
)

ggplotly(
  a %>% group_by(monthkey, ward)%>%
    summarise(total.offenses = n()) %>%
    mutate(total.offenses.pct = total.offenses/sum(total.offenses)) %>% 
    ungroup() %>% 
    mutate(ward = fct_reorder(ward, total.offenses, .fun = mean)) %>% 
    ggplot(aes(monthkey, total.offenses.pct, fill = ward)) +
    geom_area() +
    scale_y_continuous(labels = scales::percent) +
    labs(title = 'Total Offenses Percentage (%) by Month/Ward', x = '', y = '')
) 


```


# EDA: num vars

### sample data
```{r}
a %>% select(where(is.numeric)) %>% sample_n(10)
```

### glimpse structure
```{r}
a %>% select(where(is.numeric)) %>% glimpse
```

### check missing values
```{r}
a %>% select(where(is.numeric)) %>% miss_var_summary
```

# Anomaly Detection

## Total Offenses by Month
```{r warning=FALSE, out.width=1000}
library(anomalize)
# anomalize(data, target, method = c("iqr", "gesd"), alpha = 0.05, max_anoms = 0.2, verbose = FALSE)

# alpha: Controls the width of the "normal" range. Lower values are more conservative while higher values are less prone to incorrectly classifying "normal" observations.

# max_anoms: The maximum percent of anomalies permitted to be identified.

a.anomalize = a %>%
  group_by(monthkey) %>% 
  summarise(total.offenses = n()) %>%
  time_decompose(total.offenses, method = 'twitter', merge = TRUE) %>%
  anomalize(remainder, alpha = 0.30, method = 'gesd') %>%
  time_recompose()

ggplotly(
a.anomalize %>%
  filter(monthkey < as.Date('2017-11-01')) %>% 
  plot_anomalies(
    ncol = 1,
    alpha_dots = 0.5,
    alpha_circles = 0.5,
    size_circles = 1.5,
    time_recomposed = TRUE,
    alpha_ribbon = 0.05
    ) + scale_y_continuous(labels = comma) +
  labs(x = '', y = 'total.offenses', title = 'total.offenses')
)
```

<h style="color: blue; font-size:12px;">There was statistically high crime activity in August 2014</h>
<h style="color: blue; font-size:12px;">There was statistically low crime activity in August 2017</h>

# Predict Total Offenses for the 3 Months

### Create Model
```{r fig.width = 12, message=FALSE}
library(prophet)

#renaming cols to prophet's col conventions
a.agg = a %>%
  #filter(crimetype == 'Violent') %>% 
  group_by(report_dat = round_date(report_dat,'month')) %>% 
  summarise(total.offenses = n()) %>% 
  select(ds = report_dat, y = total.offenses)

#creating model
a.agg.model = a.agg %>% prophet()

#using model make future period df
a.agg.future = a.agg.model %>% make_future_dataframe(
  periods = 3, #next 2 months
  freq = 'month') 

#make forecasts df
a.agg.forecast = a.agg.model %>% predict(a.agg.future)

#plot forecast
a.agg.model %>% dyplot.prophet(a.agg.forecast)

#plot forecast components
a.agg.model %>% prophet_plot_components(a.agg.forecast)

```

<h style="color: blue; font-size:12px;">Looking at bottom chart of the component graphs, we can see crime has seasonal monthly peaks </h>

## Prediction Summary
```{r}
a.agg.forecast %>% tail(3) %>% select(ds, yhat) %>% 
  rename(month = ds, total_offenses_prediction = yhat)
```

```{r}

```

